\documentclass[twosided]{memoir}
\usepackage{mybook}
\begin{document}
\chapter*{Typesetter Comments}
I purchased this book at a small bookstore in Ottawa while on a short vacation from college. Initially, it was to read for my own knowledge and benefit. After reading the first few sections of the book, I was very engaged with Osgood's explanations and approach to complex variables (something I was rather unfamiliar with at the time). I attempted to locate copies of this book online and came up short. There appears to only be copies in libraries (or bookstores), and even that is limited; certainly there is no online or pdf version. 

Therefore, I've decided to type this text into LaTeX to the best of my ability. Being very comfortable with LaTeX already, but no expert in it (or Tikz), I decided that the purpose would be twofold: to learn how to type a book in LaTeX, complete with drawings in Tikz, and to make the textbook more available to second or third year undergraduates. With the paper in my copy becoming brittle and some of the text fading away, I wanted to preserve Osgood's work. 

There are a few modifications I made. Almost all of the text is verbatim from the book, with the majority of the changes being in the style of mathematics. For instance, $\lim\limits_{n=\infty} $ became $\lim\limits_{n \to \infty} $ and $z \text{ in } S$ became $z\in S$ to fit a modern ``syntax'' of mathematics. 



I have decided to make the source code available to the public, which can be found on my GitHub. 

\end{document}
